Tagging

### 2007-06-11 slides/data.tex
*****************************   Changed nltk_lite to nltk *****************************
This has been done in preparation for 1.0,
in which the '-Lite' part of the name will
go away.  In the meantime the distributions
will still be called NLTK-Lite, but the package
itself will be just 'nltk'.  Code, demos, tests,
tutorials, contrib are all updated.


svn/trunk@4647

\small
\begin{verbatim}
>>> from nltk_lite.corpora import toolbox
>>> lexicon = toolbox.parse_corpus('rotokas.dic', key='lx')
\end{verbatim}
\end{frame}

@@

\begin{verbatim}
>>> from nltk_lite.etree.ElementTree import SubElement
>>> for entry in lexicon:
...     for field in entry:
...	    if field.tag == 'lx':
...         cv_field = SubElement(entry, 'cv')
...         cv_field.text = cv(field.text)
\end{verbatim}
\end{frame}

@@

\tiny
\begin{verbatim}
>>> import sys
>>> from nltk_lite.etree.ElementTree import ElementTree
>>> tree = ElementTree(lexicon[3])
>>> tree.write(sys.stdout)
<record>
  <lx>kaakaaro</lx>
  <ps>N.N</ps>
  <ge>mixtures</ge>
  <gp>???</gp>
  <eng>mixtures</eng>
  <eng>charm used to keep married men and women youthful and attractive</eng>
  <cmt>Check vowel length. Is it kaakaaro or kaakaro?</cmt>
  <dt>14/May/2005</dt>
  <ex>Kaakaroto ira purapaiveira aue iava opita, voeao-pa airepa oraouirara, ra va aiopaive.</ex>
  <xp>Kokonas ol i save wokim long ol kain samting bilong ol nupela marit, bai ol i ken kaikai.</xp>
  <xe>Mixtures are made from coconut, ???.</xe>
</record>
\end{verbatim}
\end{frame}

@@

\scriptsize
\begin{verbatim}
>>> from nltk_lite.tokenize import regexp
>>> from nltk_lite.probability import FreqDist
>>> fd = FreqDist()
>>> lexemes = [lexeme.text.lower() for lexeme in lexicon.findall('record/lx')]
>>> for lex in lexemes:
...     for syl in regexp(lex, pattern=r'[^aeiou][aeiou]'):
...         fd.inc(syl)
\end{verbatim}
\end{frame}

@@

\begin{verbatim}
>>> from nltk_lite.utilities import MinimalSet
>>> pos = 1
>>> ms = MinimalSet((lex[:pos] + '_' + lex[pos+1:], lex[pos], lex)
...                 for lex in lexemes if len(lex) == 4)
>>> for context in ms.contexts(3):
...     print context + ':',
...     for target in ms.targets():
...         print "%-4s" % ms.display(context, target, "-"),
...     print
k_si: kasi -    kesi -    kosi
k_ru: karu kiru keru kuru koru
k_pu: kapu kipu -    -    kopu
k_ro: karo kiro -    -    koro
k_ri: kari kiri keri kuri kori
k_pa: kapa -    kepa -    kopa
k_ra: kara kira kera -    kora
k_ku: kaku -    -    kuku koku
k_ki: kaki kiki -    -    koki
\end{verbatim}
\end{frame}

@@

\begin{frame}[fragile]
\frametitle{Sorting and Lookup}
\scriptsize
\begin{verbatim}
>>> from nltk_lite.utilities import edit_dist
>>> def fuzzy_spell(target):
...     scored_candidates = []
...     code = soundex(target)
...     for word in soundex_idx[code]:
...         dist = edit_dist(word, target)
...         scored_candidates.append((dist, word))
...     scored_candidates.sort()
...     return [w for (d,w) in scored_candidates[:10]]
\end{verbatim}

@@

\begin{verbatim}
>>> from nltk_lite import parse
>>> grammar = parse.cfg.parse_grammar('''
...   S -> Head "ps" Glosses Comment "dt" Examples
...   Head -> "lx" | "lx" "rt"
...   Glosses -> Gloss Glosses
...   Glosses ->
...   Gloss -> "ge" | "gp"
...   Examples -> Example Examples
...   Examples ->
...   Example -> "ex" "xp" "xe"
...   Comment -> "cmt"
...   Comment ->
...   ''')
>>> rd_parser = parse.RecursiveDescent(grammar)
\end{verbatim}
\end{frame}

>>>>

\small
\begin{verbatim}
>>> from nltk.corpora import toolbox
>>> lexicon = toolbox.parse_corpus('rotokas.dic', key='lx')
\end{verbatim}
\end{frame}

@@

\begin{verbatim}
>>> from nltk.etree.ElementTree import SubElement
>>> for entry in lexicon:
...     for field in entry:
...	    if field.tag == 'lx':
...         cv_field = SubElement(entry, 'cv')
...         cv_field.text = cv(field.text)
\end{verbatim}
\end{frame}

@@

\tiny
\begin{verbatim}
>>> import sys
>>> from nltk.etree.ElementTree import ElementTree
>>> tree = ElementTree(lexicon[3])
>>> tree.write(sys.stdout)
<record>
  <lx>kaakaaro</lx>
  <ps>N.N</ps>
  <ge>mixtures</ge>
  <gp>???</gp>
  <eng>mixtures</eng>
  <eng>charm used to keep married men and women youthful and attractive</eng>
  <cmt>Check vowel length. Is it kaakaaro or kaakaro?</cmt>
  <dt>14/May/2005</dt>
  <ex>Kaakaroto ira purapaiveira aue iava opita, voeao-pa airepa oraouirara, ra va aiopaive.</ex>
  <xp>Kokonas ol i save wokim long ol kain samting bilong ol nupela marit, bai ol i ken kaikai.</xp>
  <xe>Mixtures are made from coconut, ???.</xe>
</record>
\end{verbatim}
\end{frame}

@@

\scriptsize
\begin{verbatim}
>>> from nltk.tokenize import regexp
>>> from nltk.probability import FreqDist
>>> fd = FreqDist()
>>> lexemes = [lexeme.text.lower() for lexeme in lexicon.findall('record/lx')]
>>> for lex in lexemes:
...     for syl in regexp(lex, pattern=r'[^aeiou][aeiou]'):
...         fd.inc(syl)
\end{verbatim}
\end{frame}

@@

\begin{verbatim}
>>> from nltk.utilities import MinimalSet
>>> pos = 1
>>> ms = MinimalSet((lex[:pos] + '_' + lex[pos+1:], lex[pos], lex)
...                 for lex in lexemes if len(lex) == 4)
>>> for context in ms.contexts(3):
...     print context + ':',
...     for target in ms.targets():
...         print "%-4s" % ms.display(context, target, "-"),
...     print
k_si: kasi -    kesi -    kosi
k_ru: karu kiru keru kuru koru
k_pu: kapu kipu -    -    kopu
k_ro: karo kiro -    -    koro
k_ri: kari kiri keri kuri kori
k_pa: kapa -    kepa -    kopa
k_ra: kara kira kera -    kora
k_ku: kaku -    -    kuku koku
k_ki: kaki kiki -    -    koki
\end{verbatim}
\end{frame}

@@

\begin{frame}[fragile]
\frametitle{Sorting and Lookup}
\scriptsize
\begin{verbatim}
>>> from nltk.utilities import edit_dist
>>> def fuzzy_spell(target):
...     scored_candidates = []
...     code = soundex(target)
...     for word in soundex_idx[code]:
...         dist = edit_dist(word, target)
...         scored_candidates.append((dist, word))
...     scored_candidates.sort()
...     return [w for (d,w) in scored_candidates[:10]]
\end{verbatim}

@@

\begin{verbatim}
>>> from nltk import parse
>>> grammar = parse.cfg.parse_grammar('''
...   S -> Head "ps" Glosses Comment "dt" Examples
...   Head -> "lx" | "lx" "rt"
...   Glosses -> Gloss Glosses
...   Glosses ->
...   Gloss -> "ge" | "gp"
...   Examples -> Example Examples
...   Examples ->
...   Example -> "ex" "xp" "xe"
...   Comment -> "cmt"
...   Comment ->
...   ''')
>>> rd_parser = parse.RecursiveDescent(grammar)
\end{verbatim}
\end{frame}

### 2007-06-11 slides/introduction.tex
*****************************   Changed nltk_lite to nltk *****************************
This has been done in preparation for 1.0,
in which the '-Lite' part of the name will
go away.  In the meantime the distributions
will still be called NLTK-Lite, but the package
itself will be just 'nltk'.  Code, demos, tests,
tutorials, contrib are all updated.


svn/trunk@4647

\begin{verbatim}
    >>> from nltk_lite.corpora import cmudict
    >>> from string import join
    >>> for word, num, pron in cmudict.raw():
    ...     stress_pattern = join(c for c in join(pron) if c in "012")
    ...     if stress_pattern.endswith("1 0 0 0 0"):
    ...         print word, "/", join(pron)
    ACCUMULATIVELY / AH0 K Y UW1 M Y AH0 L AH0 T IH0 V L IY0
    AGONIZINGLY / AE1 G AH0 N AY0 Z IH0 NG L IY0
    CARICATURIST / K EH1 R AH0 K AH0 CH ER0 AH0 S T
    CIARAMITARO / CH ER1 AA0 M IY0 T AA0 R OW0
    CUMULATIVELY / K Y UW1 M Y AH0 L AH0 T IH0 V L IY0
    DEBENEDICTIS / D EH1 B EH0 N AH0 D IH0 K T AH0 S
    DELEONARDIS / D EH1 L IY0 AH0 N AA0 R D AH0 S
    FORMALIZATION / F AO1 R M AH0 L AH0 Z EY0 SH AH0 N
    GIANNATTASIO / JH AA1 N AA0 T AA0 S IY0 OW0
    HYPERSENSITIVITY / HH AY2 P ER0 S EH1 N S AH0 T IH0 V AH0 T IY0
    IMAGINATIVELY / IH2 M AE1 JH AH0 N AH0 T IH0 V L IY0
    INSTITUTIONALIZES / IH2 N S T AH0 T UW1 SH AH0 N AH0 L AY0 Z AH0 Z
    INSTITUTIONALIZING / IH2 N S T AH0 T UW1 SH AH0 N AH0 L AY0 Z IH0 NG
    MANGIARACINA / M AA1 N JH ER0 AA0 CH IY0 N AH0
    SPIRITUALIST / S P IH1 R IH0 CH AH0 W AH0 L AH0 S T
    SPIRITUALISTS / S P IH1 R IH0 CH AH0 W AH0 L AH0 S T S
    SPIRITUALISTS / S P IH1 R IH0 CH AH0 W AH0 L AH0 S S
    SPIRITUALISTS / S P IH1 R IH0 CH AH0 W AH0 L AH0 S
    SPIRITUALLY / S P IH1 R IH0 CH AH0 W AH0 L IY0
    UNALIENABLE / AH0 N EY1 L IY0 EH0 N AH0 B AH0 L
    UNDERKOFFLER / AH1 N D ER0 K AH0 F AH0 L ER0
\end{verbatim}
\end{frame}

\subsection{Minimal Sets from Lexicon}
\begin{frame}[fragile]
  \frametitle{Minimal Sets from Lexicon}
  \tiny
\begin{verbatim}
    >>> from nltk_lite.corpora import shoebox
    >>> from nltk_lite.utilities import MinimalSet
    >>> length, position, min = 4, 1, 3
    >>> lexemes = [field[1].lower() for entry in shoebox.raw('rotokas')
    ...		   for field in entry if field[0] == 'lx']
    >>> ms = MinimalSet()
    >>> for lex in lexemes:
    ...     if len(lex) == length:
    ...         context = lex[:position] + '_' + lex[position+1:]
    ...         target = lex[position]
    ...         ms.add(context, target, lex)
    >>> for context in ms.contexts(3):
    ...     for target in ms.targets():
    ...         print "%-4s" % ms.display(context, target, "-"),
    ...     print
    kasi -    kesi kusi kosi
    kava -    -    kuva kova
    karu kiru keru kuru koru
    kapu kipu -    -    kopu
    karo kiro -    -    koro
    kari kiri keri kuri kori
    kapa -    kepa -    kopa
    kara kira kera -    kora
    kaku -    -    kuku koku
    kaki kiki -    -    koki
\end{verbatim}
\end{frame}

\subsection{Modelling Text Genres}
\begin{frame}[fragile]
  \frametitle{Modelling Text Genres}
  \tiny
\begin{verbatim}
    >>> from nltk_lite.corpora import genesis
    >>> from nltk_lite.probability import ConditionalFreqDist
    >>> from nltk_lite.utilities import print_string
    >>> cfdist = ConditionalFreqDist()
    >>> prev = None
    >>> for word in genesis.raw():
    ...     word = word.lower()
    ...     cfdist[prev].inc(word)
    ...     prev = word
    >>> words = []
    >>> prev = 'lo,'
    >>> for i in range(99):
    ...     words.append(prev)
    ...     for word in cfdist[prev].sorted_samples():
    ...         if word not in words:
    ...             break
    ...     prev = word
    >>> print_string(join(words))
    lo, it came to the land of his father and he said, i will not be a
    wife unto him, saying, if thou shalt take our money in their kind,
    cattle, in thy seed after these are my son from off any more than all
    that is this day with him into egypt, he, hath taken away unawares to
    pass, when she bare jacob said one night, because they were born two
    hundred years old, as for an altar there, he had made me out at her
    pitcher upon every living creature after thee shall come near her:
    yea,
\end{verbatim}
\end{frame}

\subsection{Exploring Syntax}
\begin{frame}[fragile]
  \frametitle{Exploring Syntax}
  \tiny
\begin{verbatim}
    >>> from nltk_lite.corpora import treebank
    >>> from string import join
    >>> def vp_conj(tree):
    ...     if tree.node == 'VP' and len(tree) == 3 and tree[1].leaves() == ['but']:
    ...         return True
    ...     else:
    ...         return False
    >>> for tree in treebank.parsed():
    ...     for vp1,conj,vp2 in tree.subtrees(vp_conj):
    ...         print join(child.node for child in vp1), "*BUT*", join(child.node for child in vp2)
    VBP ADVP-TMP PP-PRD PP *BUT* VBP VP
    VBZ VP *BUT* VBZ NP PP-CLR
    PP-TMP VBZ VP *BUT* VBD ADVP-TMP S
    VBZ SBAR *BUT* VBZ SBAR
    VBD SBAR *BUT* VBD RB VP
    VBD SBAR *BUT* VBD S
    VBP NP-PRD *BUT* VBP RB ADVP-TMP VP
    VBN PP PP-TMP *BUT* ADVP-TMP VBN NP
    MD VP *BUT* VBZ NP SBAR-ADV
    VBD ADVP-CLR *BUT* VBD NP
    VBN NP PP *BUT* VBN NP PP SBAR-PRP
    VBD NP *BUT* MD RB VP
    VBD NP PP-CLR *BUT* VBD PRT NP
    VBZ S *BUT* MD VP
\end{verbatim}
\end{frame}

@@

\begin{frame}[fragile]
\frametitle{Installing Python and NLTK}
\begin{enumerate}
\item Install Python, Numeric
\item Install NLTK-Lite, NLTK-Lite-Corpora
\item Optional: Matplotlib, WordNet
\item Set environment variable \verb|NLTK_LITE_CORPORA|
\end{enumerate}

>>>>

\begin{verbatim}
    >>> from nltk.corpora import cmudict
    >>> from string import join
    >>> for word, num, pron in cmudict.raw():
    ...     stress_pattern = join(c for c in join(pron) if c in "012")
    ...     if stress_pattern.endswith("1 0 0 0 0"):
    ...         print word, "/", join(pron)
    ACCUMULATIVELY / AH0 K Y UW1 M Y AH0 L AH0 T IH0 V L IY0
    AGONIZINGLY / AE1 G AH0 N AY0 Z IH0 NG L IY0
    CARICATURIST / K EH1 R AH0 K AH0 CH ER0 AH0 S T
    CIARAMITARO / CH ER1 AA0 M IY0 T AA0 R OW0
    CUMULATIVELY / K Y UW1 M Y AH0 L AH0 T IH0 V L IY0
    DEBENEDICTIS / D EH1 B EH0 N AH0 D IH0 K T AH0 S
    DELEONARDIS / D EH1 L IY0 AH0 N AA0 R D AH0 S
    FORMALIZATION / F AO1 R M AH0 L AH0 Z EY0 SH AH0 N
    GIANNATTASIO / JH AA1 N AA0 T AA0 S IY0 OW0
    HYPERSENSITIVITY / HH AY2 P ER0 S EH1 N S AH0 T IH0 V AH0 T IY0
    IMAGINATIVELY / IH2 M AE1 JH AH0 N AH0 T IH0 V L IY0
    INSTITUTIONALIZES / IH2 N S T AH0 T UW1 SH AH0 N AH0 L AY0 Z AH0 Z
    INSTITUTIONALIZING / IH2 N S T AH0 T UW1 SH AH0 N AH0 L AY0 Z IH0 NG
    MANGIARACINA / M AA1 N JH ER0 AA0 CH IY0 N AH0
    SPIRITUALIST / S P IH1 R IH0 CH AH0 W AH0 L AH0 S T
    SPIRITUALISTS / S P IH1 R IH0 CH AH0 W AH0 L AH0 S T S
    SPIRITUALISTS / S P IH1 R IH0 CH AH0 W AH0 L AH0 S S
    SPIRITUALISTS / S P IH1 R IH0 CH AH0 W AH0 L AH0 S
    SPIRITUALLY / S P IH1 R IH0 CH AH0 W AH0 L IY0
    UNALIENABLE / AH0 N EY1 L IY0 EH0 N AH0 B AH0 L
    UNDERKOFFLER / AH1 N D ER0 K AH0 F AH0 L ER0
\end{verbatim}
\end{frame}

\subsection{Minimal Sets from Lexicon}
\begin{frame}[fragile]
  \frametitle{Minimal Sets from Lexicon}
  \tiny
\begin{verbatim}
    >>> from nltk.corpora import shoebox
    >>> from nltk.utilities import MinimalSet
    >>> length, position, min = 4, 1, 3
    >>> lexemes = [field[1].lower() for entry in shoebox.raw('rotokas')
    ...		   for field in entry if field[0] == 'lx']
    >>> ms = MinimalSet()
    >>> for lex in lexemes:
    ...     if len(lex) == length:
    ...         context = lex[:position] + '_' + lex[position+1:]
    ...         target = lex[position]
    ...         ms.add(context, target, lex)
    >>> for context in ms.contexts(3):
    ...     for target in ms.targets():
    ...         print "%-4s" % ms.display(context, target, "-"),
    ...     print
    kasi -    kesi kusi kosi
    kava -    -    kuva kova
    karu kiru keru kuru koru
    kapu kipu -    -    kopu
    karo kiro -    -    koro
    kari kiri keri kuri kori
    kapa -    kepa -    kopa
    kara kira kera -    kora
    kaku -    -    kuku koku
    kaki kiki -    -    koki
\end{verbatim}
\end{frame}

\subsection{Modelling Text Genres}
\begin{frame}[fragile]
  \frametitle{Modelling Text Genres}
  \tiny
\begin{verbatim}
    >>> from nltk.corpora import genesis
    >>> from nltk.probability import ConditionalFreqDist
    >>> from nltk.utilities import print_string
    >>> cfdist = ConditionalFreqDist()
    >>> prev = None
    >>> for word in genesis.raw():
    ...     word = word.lower()
    ...     cfdist[prev].inc(word)
    ...     prev = word
    >>> words = []
    >>> prev = 'lo,'
    >>> for i in range(99):
    ...     words.append(prev)
    ...     for word in cfdist[prev].sorted_samples():
    ...         if word not in words:
    ...             break
    ...     prev = word
    >>> print_string(join(words))
    lo, it came to the land of his father and he said, i will not be a
    wife unto him, saying, if thou shalt take our money in their kind,
    cattle, in thy seed after these are my son from off any more than all
    that is this day with him into egypt, he, hath taken away unawares to
    pass, when she bare jacob said one night, because they were born two
    hundred years old, as for an altar there, he had made me out at her
    pitcher upon every living creature after thee shall come near her:
    yea,
\end{verbatim}
\end{frame}

\subsection{Exploring Syntax}
\begin{frame}[fragile]
  \frametitle{Exploring Syntax}
  \tiny
\begin{verbatim}
    >>> from nltk.corpora import treebank
    >>> from string import join
    >>> def vp_conj(tree):
    ...     if tree.node == 'VP' and len(tree) == 3 and tree[1].leaves() == ['but']:
    ...         return True
    ...     else:
    ...         return False
    >>> for tree in treebank.parsed():
    ...     for vp1,conj,vp2 in tree.subtrees(vp_conj):
    ...         print join(child.node for child in vp1), "*BUT*", join(child.node for child in vp2)
    VBP ADVP-TMP PP-PRD PP *BUT* VBP VP
    VBZ VP *BUT* VBZ NP PP-CLR
    PP-TMP VBZ VP *BUT* VBD ADVP-TMP S
    VBZ SBAR *BUT* VBZ SBAR
    VBD SBAR *BUT* VBD RB VP
    VBD SBAR *BUT* VBD S
    VBP NP-PRD *BUT* VBP RB ADVP-TMP VP
    VBN PP PP-TMP *BUT* ADVP-TMP VBN NP
    MD VP *BUT* VBZ NP SBAR-ADV
    VBD ADVP-CLR *BUT* VBD NP
    VBN NP PP *BUT* VBN NP PP SBAR-PRP
    VBD NP *BUT* MD RB VP
    VBD NP PP-CLR *BUT* VBD PRT NP
    VBZ S *BUT* MD VP
\end{verbatim}
\end{frame}

@@

\begin{frame}[fragile]
\frametitle{Installing Python and NLTK}
\begin{enumerate}
\item Install Python, Install NLTK, NLTK-Corpora
\item Optional: Numeric, Matplotlib
\item Set environment variable \verb|NLTK_CORPORA|
\end{enumerate}

### 2007-06-11 slides/programming.tex
*****************************   Changed nltk_lite to nltk *****************************
This has been done in preparation for 1.0,
in which the '-Lite' part of the name will
go away.  In the meantime the distributions
will still be called NLTK-Lite, but the package
itself will be just 'nltk'.  Code, demos, tests,
tutorials, contrib are all updated.


svn/trunk@4647

\begin{verbatim}
  >>> from nltk_lite.corpora import gutenberg
  >>> count = {}
  >>> for word in gutenberg.raw('shakespeare-macbeth'):
  ...     word = word.lower()
  ...     if word not in count:
  ...         count[word] = 0
  ...     count[word] += 1
\end{verbatim}

@@

\begin{verbatim}
  >>> import re
  >>> from nltk_lite.utilities import re_show
  >>> sent = "colourless green ideas sleep furiously"
\end{verbatim}

@@

\begin{verbatim}
  >>> from nltk_lite import utilities
  >>> utilities.re_show('green', s)
  colourless {green} ideas sleep furiously
\end{verbatim}
\end{itemize}
\end{frame}

@@

\begin{verbatim}
  >>> from nltk_lite.corpora import gutenberg
  >>> gutenberg.items
  ['austen-emma', 'austen-persuasion', 'austen-sense', 'bible-kjv', 'blake-poems', 'blake-songs', 'chesterton-ball', 'chesterton-brown', 'chesterton-thursday', 'milton-paradise', 'shakespeare-caesar', 'shakespeare-hamlet', 'shakespeare-macbeth', 'whitman-leaves']
  >>> count = 0
  >>> for word in gutenberg.raw('whitman-leaves'):
  ...     count += 1
  >>> print count
  154873
\end{verbatim}
\end{frame}

@@

\begin{verbatim}
  >>> from nltk_lite.corpora import brown
  >>> brown.items
  ['a', 'b', 'c', 'd', 'e', 'f', 'g', 'h', 'j', 'k', 'l', 'm', 'n', 'p', 'r']
  >>> from nltk_lite.corpora import extract
  >>> print extract(0, brown.raw())
  ['The', 'Fulton', 'County', 'Grand', 'Jury', 'said', 'Friday', 'an', 'investigation', 'of', "Atlanta's", 'recent', 'primary', 'election', 'produced', '``', 'no', 'evidence', "''", 'that', 'any', 'irregularities', 'took', 'place', '.']
  >>> print extract(0, brown.tagged())
  [('The', 'at'), ('Fulton', 'np-tl'), ('County', 'nn-tl'), ('Grand', 'jj-tl'), ('Jury', 'nn-tl'), ('said', 'vbd'), ('Friday', 'nr'), ('an', 'at'), ('investigation', 'nn'), ('of', 'in'), ("Atlanta's", 'np$'), ('recent', 'jj'), ('primary', 'nn'), ('election', 'nn'), ('produced', 'vbd'), ('``', '``'), ('no', 'at'), ('evidence', 'nn'), ("''", "''"), ('that', 'cs'), ('any', 'dti'), ('irregularities', 'nns'), ('took', 'vbd'), ('place', 'nn'), ('.', '.')]
\end{verbatim}
\end{frame}

@@

\begin{verbatim}
  >>> from nltk_lite.corpora import treebank
  >>> print extract(0, treebank.parsed())
  (S:
    (NP-SBJ:
      (NP: (NNP: 'Pierre') (NNP: 'Vinken'))
      (,: ',')
      (ADJP: (NP: (CD: '61') (NNS: 'years')) (JJ: 'old'))
      (,: ','))
    (VP:
      (MD: 'will')
      (VP:
        (VB: 'join')
        (NP: (DT: 'the') (NN: 'board'))
        (PP-CLR:
          (IN: 'as')
          (NP: (DT: 'a') (JJ: 'nonexecutive') (NN: 'director')))
        (NP-TMP: (NNP: 'Nov.') (CD: '29'))))
    (.: '.'))
\end{verbatim}
\end{frame}

>>>>

\begin{verbatim}
  >>> from nltk.corpora import gutenberg
  >>> count = {}
  >>> for word in gutenberg.raw('shakespeare-macbeth'):
  ...     word = word.lower()
  ...     if word not in count:
  ...         count[word] = 0
  ...     count[word] += 1
\end{verbatim}

@@

\begin{verbatim}
  >>> import re
  >>> from nltk.utilities import re_show
  >>> sent = "colourless green ideas sleep furiously"
\end{verbatim}

@@

\begin{verbatim}
  >>> from nltk import utilities
  >>> utilities.re_show('green', s)
  colourless {green} ideas sleep furiously
\end{verbatim}
\end{itemize}
\end{frame}

@@

\begin{verbatim}
  >>> from nltk.corpora import gutenberg
  >>> gutenberg.items
  ['austen-emma', 'austen-persuasion', 'austen-sense', 'bible-kjv', 'blake-poems', 'blake-songs', 'chesterton-ball', 'chesterton-brown', 'chesterton-thursday', 'milton-paradise', 'shakespeare-caesar', 'shakespeare-hamlet', 'shakespeare-macbeth', 'whitman-leaves']
  >>> count = 0
  >>> for word in gutenberg.raw('whitman-leaves'):
  ...     count += 1
  >>> print count
  154873
\end{verbatim}
\end{frame}

@@

\begin{verbatim}
  >>> from nltk.corpora import brown
  >>> brown.items
  ['a', 'b', 'c', 'd', 'e', 'f', 'g', 'h', 'j', 'k', 'l', 'm', 'n', 'p', 'r']
  >>> from nltk.corpora import extract
  >>> print extract(0, brown.raw())
  ['The', 'Fulton', 'County', 'Grand', 'Jury', 'said', 'Friday', 'an', 'investigation', 'of', "Atlanta's", 'recent', 'primary', 'election', 'produced', '``', 'no', 'evidence', "''", 'that', 'any', 'irregularities', 'took', 'place', '.']
  >>> print extract(0, brown.tagged())
  [('The', 'at'), ('Fulton', 'np-tl'), ('County', 'nn-tl'), ('Grand', 'jj-tl'), ('Jury', 'nn-tl'), ('said', 'vbd'), ('Friday', 'nr'), ('an', 'at'), ('investigation', 'nn'), ('of', 'in'), ("Atlanta's", 'np$'), ('recent', 'jj'), ('primary', 'nn'), ('election', 'nn'), ('produced', 'vbd'), ('``', '``'), ('no', 'at'), ('evidence', 'nn'), ("''", "''"), ('that', 'cs'), ('any', 'dti'), ('irregularities', 'nns'), ('took', 'vbd'), ('place', 'nn'), ('.', '.')]
\end{verbatim}
\end{frame}

@@

\begin{verbatim}
  >>> from nltk.corpora import treebank
  >>> print extract(0, treebank.parsed())
  (S:
    (NP-SBJ:
      (NP: (NNP: 'Pierre') (NNP: 'Vinken'))
      (,: ',')
      (ADJP: (NP: (CD: '61') (NNS: 'years')) (JJ: 'old'))
      (,: ','))
    (VP:
      (MD: 'will')
      (VP:
        (VB: 'join')
        (NP: (DT: 'the') (NN: 'board'))
        (PP-CLR:
          (IN: 'as')
          (NP: (DT: 'a') (JJ: 'nonexecutive') (NN: 'director')))
        (NP-TMP: (NNP: 'Nov.') (CD: '29'))))
    (.: '.'))
\end{verbatim}
\end{frame}

### 2007-06-11 slides/tag.tex
*****************************   Changed nltk_lite to nltk *****************************
This has been done in preparation for 1.0,
in which the '-Lite' part of the name will
go away.  In the meantime the distributions
will still be called NLTK-Lite, but the package
itself will be just 'nltk'.  Code, demos, tests,
tutorials, contrib are all updated.


svn/trunk@4647

\begin{frame}[fragile]
  \frametitle{Tagging in NLTK}
\textcolor{black}{The simplest possible tagger tags everything as a noun:}
\small
\begin{verbatim}
from nltk_lite import tokenize
text = 'There are 11 players in a football team'
text_tokens = list(tokenize.whitespace(text))
# ['There', 'are', '11', 'players', 'in', 'a', 'football', 'team']
\end{verbatim}
  \pause
\begin{verbatim}
from nltk_lite import tag
mytagger = tag.Default('nn')
for t in mytagger.tag(text_tokens):
    print t
# ('There', 'NN')
# ('are', 'NN')
# ...
\end{verbatim}
\end{frame}

@@

\pause
Training a UnigramTagger on the Penn Treebank:
{\small
\begin{verbatim}
from nltk_lite.corpora import treebank
from itertools import islice

@@

  \begin{itemize}
  \item Does not find the most likely sequence of tags, simply works
    left to right always assigning the most probable single tag (given
    the previous tag assignments)
  \item Does not cope with zero probability problem well (no smoothing
    or discounting)
  \item see module \verb|nltk_lite.tag.hmm|
  \end{itemize}
\end{frame}

>>>>

\begin{frame}[fragile]
  \frametitle{Tagging in NLTK}
\textcolor{black}{The simplest possible tagger tags everything as a noun:}
\small
\begin{verbatim}
from nltk import tokenize
text = 'There are 11 players in a football team'
text_tokens = list(tokenize.whitespace(text))
# ['There', 'are', '11', 'players', 'in', 'a', 'football', 'team']
\end{verbatim}
  \pause
\begin{verbatim}
from nltk import tag
mytagger = tag.Default('nn')
for t in mytagger.tag(text_tokens):
    print t
# ('There', 'NN')
# ('are', 'NN')
# ...
\end{verbatim}
\end{frame}

@@

\pause
Training a UnigramTagger on the Penn Treebank:
{\small
\begin{verbatim}
from nltk.corpora import treebank
from itertools import islice

@@

  \begin{itemize}
  \item Does not find the most likely sequence of tags, simply works
    left to right always assigning the most probable single tag (given
    the previous tag assignments)
  \item Does not cope with zero probability problem well (no smoothing
    or discounting)
  \item see module \verb|nltk.tag.hmm|
  \end{itemize}
\end{frame}
